---
title: "Inverse Tranform Sampling Method"
author: "Josh Soboil"
date: "21/11/2020"
output: powerpoint_presentation
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
set.seed(1000)
```

#  Inverse Transform Method

Random samples are drawn from uniform distribution, i.e. $x\sim  U(0, 1)$. For instance
```{r sample from uniform distribution, echo = TRUE}
x <- runif(n = 10000)
```

```{r plot samples, echo = TRUE}
ggplot() + 
 geom_histogram(aes(x), binwidth = 0.05, colour = "black", fill = "white", alpha = 0.95) + 
 ylab(label = "Sample frequency") +
 xlab(label = expression(paste(italic(Uniform), " ~ " (0, 1)))) +
 theme_linedraw(base_family = "Times New Roman", base_size = 12)
```
For example, from here, it is then possible to transform these samples into the inverse cdf (quantile function) of the exponential distribution and normal distribution 
```{r quantile function, echo = TRUE, warning = FALSE}
inv_cdf_norm <- qnorm(p = x, mean = 10, sd = sqrt(0.5))
```
The result of this can  be used to inform the probability for a range of parameter values occurring.
```{r plot pdfs, echo = TRUE, warning = FALSE}
ggplot() + 
 geom_density(aes(x = inv_cdf_norm), fill = "skyblue", alpha = 0.65) +
 ylab(label = "Probability Density") +
 xlab(label = "Parameter values") +
 xlim(7, 13) +
 theme_linedraw(base_family = "Times New Roman", base_size = 12) +
 stat_ecdf(aes(inv_cdf_norm), colour = "red", alpha = 0.5, lwd = 0.95) +
 theme_linedraw(base_family = "Times New Roman", base_size = 12) +
 annotate("segment", x = 8, xend = 9.29, y = 0.5, yend = 0.35, colour = "black", 
          size = .4, alpha = 0.75, arrow = arrow()) +
 annotate("text", x = 8, y = .54, label = "pdf", 
          size = 5, family = "Times New Roman") +
 annotate("segment", x = 12, xend = 10.4, y = 0.5, yend = 0.7, colour = "black", 
          size = .4, alpha = 0.75, arrow = arrow()) +
 annotate("text", x = 12, y = .46, label = "cdf", 
          size = 5, family = "Times New Roman")
``` %>% 