---
title: "Dutch Book proof"
output:
 html_document:
  theme: readable
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(matlib)
```

##  An incoherent (Dutch Book) series of bets:
Consider a hypothetical event $\theta$ and its set of possible outcomes. where $\theta = 1$ indicates that the event occurred and $\theta = 0$ indicates that it did not. We can subsequently define a bet to be worth an associated stake $S$ if $\theta$ does occur, and nothing if $\theta$ does not occur, i.e. we can either win our bet with a total payout $S$ or lose. Furthermore, we assume that the event is strictly binary, i.e. mutually exclusive. 

A bookmaker is typically known to sell a bet at a price $\pi_{\theta}S$. Likewise, the ratio $\pi:(1-\pi)$ is the betting odds in favour of the event $\theta$. We can then denote the action of placing a bet on $\theta$ as $a_{S,\theta}$. From this, we are able to also define the consequences of the action. For instance, if $\theta=1$, the purchaser of the bet will have a net gain of $(1-\pi)S$, which is simply the $S$ minus the initial buy-in price $\pi S$; similarly, if $\theta = 0$, a net gain of $-\pi S$.

The rules for these gains can notationally represented as
$$
g_{\theta} = \left\{
    \begin{array}{ll}
        (1-\pi_{i})S &\mbox{if } \theta=1\\
        -\pi_{i}S &\mbox{if } \theta=0
    \end{array}
\right.
$$
However, this is from the perspective of a gambler and not a bookie. In order to fully comprehend this scenario, however, we have to invert this function. The reason for this is that, in reality, we are the bookie and, much like the bookie, we have our own probability judgements of an event's outcomes $\pi$, and associated stakes $S$ thereof. This is consequently sometimes described as being a bookmaker with 'nature'. Nevertheless, below is the same gain function $g_\theta$ inverted.
$$
g_{\theta} = \left\{
    \begin{array}{ll}
        -(1-\pi_{i})S &\mbox{if } \theta=1\\
          \pi_{i}S &\mbox{if } \theta=0
    \end{array}
\right.
$$

Now consider a more complex scenario, where we place a series of separate bets on $3$ mutually exclusive events, $\theta_{1}$, $\theta_{2}$, and $\theta_3$, each with a corresponding stake $S_\theta$, forming a vector $s=(S_{\theta_{1}}, S_{\theta_{2}}, S_{\theta_{3}})$. We likewise consider the probability of each event $\theta$, and assign the following probabilities $\pi_{\theta_{1}}$, $\pi_{\theta_{2}}$, and $\pi_{\theta_{3}}$,
```{r}
pi1 <- .5
pi2 <- .8
pi3 <- .25
```
Because the events are mutually exclusive, we can conveniently construct our probabilities in an $n\times n$ square matrix $R$ to track all the possible realised events and the outcomes therein. Intuitively, for example, this theoretically means that $1-\pi_{\theta_{1}}=\pi_{\theta_{2}}+\pi_{\theta_{3}}$, and vice versa for the other events. 
```{r}
R <- matrix(
 c(1 - pi1,   - pi2,   - pi3,
     - pi1, 1 - pi2,   - pi3,
     - pi1,   - pi2, 1 - pi3),
            nrow = 3)
R
```
Accordingly, we can use the vectors of stakes $s$ multiplied with our $n\times n$ matrix $R$  to solve for our vector gains $g_{\theta}$, such that $g_{\theta}=Rs$.

However, it is now critical to bring to light the following: our set of bets do not conform to Kolmogorov’s probability axiom $P(A\cap, ...,\cap A) = 1$ for mutually exclusive events. As the sum of the bets is
```{r, echo = TRUE}
sum(pi1 + pi2 + pi3)
```
which clearly means that $\pi_{\theta_{1}}+\pi_{\theta_{2}}+\pi_{\theta_{3}}\neq1$. This, consequently, suggests that the matrix $R$ is invertible. Technically speaking, and using a more commonly adopted notation, there is a scalar $x$ that gives the desired output $b = Ax$. As a corollary, this also means that the determinant of the matrix $R$ is $|{R}| > 0$, as evidence by
```{r, echo = TRUE}
det(R)
```
It is blatantly important to prove whether our matrix $R$ does conform to this conditions, where $I = R^{-1}R$. However, we must initially digress by explaining the properties of an identity matrix $I$ for those of us who are unfamiliar. 

An identity matrix $I$ is where all the diagonals of a matrix are $1$'s, with the rest of the elements only containing $0$'s. It is, broadly speaking, a convenient tool used for matrix manipulations. Below is an example of what a $n\times n$ identity matrix $I$ will look like:
```{r, echo = FALSE}
I <- diag(c(1, 1, 1))
I
```
It is, therefore, a literal 'do nothing' matrix, as is clearly evidence by solving the following,
```{r, echo = TRUE}
b <- c(100, 101, 102)
I%*%b
```
Evidently, the multiplication $I\times b$ does not change the values of $b$, such that $I\times b = b$. This, then, brings us back to the focal point of this section. 

If $R$ is indeed an invertible matrix, we will be able to recreate the same identity matrix $I$ as above, if $I = R^{-1}R$.
```{r, echo = TRUE}
round(inv(R)%*%R, digits = 5)
```
The result of applying $R^{-1}R$ clearly mirrors the identity matrix example given above. As has already been alluded to, proving $I = R^{-1}R$ can be thought as a soft proof for whether any combination of linear equations is solvable if we wanted, for example, to bring $R$ to the other side of the equation. Consequently, if it is proven, this enables us to solve for a vector of stakes $s$, rather than $g_{\theta}$, by assigning $g_{\theta}$ a vector of desired gains. For instance, if we assign the vector $g_{\theta}$ the following values 
```{r, echo = TRUE}
g_theta <- c(32, -1, 41)
g_theta
```
it possible to solve for the linear transformation of $s$. First we multiply each side of the equation by $R^{-1}$, so that $R^{-1}Rs=R^{-1}g$, thus $I\times s = s =R^{-1}g$, and, ultimately $s = R^{-1}g$. Accordingly, we solve for a vector of stakes $s$ that guarantees a net positive gain:
```{r, echo = TRUE}
s_solved <- inv(R)%*%g_theta
s_solved
```
then, as is obvious,
```{r, echo = TRUE}
R%*%s_solved
```
In sum then, it is clear that this conclusively proves the fact that, if the probability matrix $R$ meets the invertible condition, such that $I = R^{-1}R$, we can solve for an unknown vector of stakes $s$ by equating it with the product of a linear combination of desired gains $g_{\theta}$ and an associated probability matrix $R$. In technical language, this means that when $R$ is an invertible matrix, the solution to $Rs=g_{\theta}$ is $s=R^{-1}g$.