---
title: "Dutch Book proof"
output:
 html_document:
  theme: readable
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(matlib)
```

##  An incoherent (Dutch Book) series of bets:
Consider a hypothetical event $\theta$ and its set of binary outcomes, where $\theta = 1$ indicates the event occurred and $\theta = 0$ indicates it did not. We can subsequently define a bet to be worth an associated stake $S$ if $\theta$ does occur, and nothing if $\theta$ does not occur, i.e. we can either win our bet with a total payout $S$ or lose.

Further consider a scenario where we are offered a bet a bookie, who is typically known to sell a bet at a price $\pi_{\theta}S$. Likewise, the ratio $\pi:(1-\pi)$ is the betting odds in favour of the event $\theta$. We can also denote the action of placing a bet on $\theta$ as $a_{S,\theta}$. This enables us to associate consequences with an action. For instance, if $\theta=1$, the purchaser of the bet will have a net gain of $(1-\pi)S$, which is simply the stake $S$ minus the initial buy-in price $\pi S$; similarly, if $\theta = 0$, a net gain of $-\pi S$. The rules for these gains can notationally represented by the function,
$$
g_{\theta} = \left\{
    \begin{array}{ll}
        (1-\pi_{i})S &\mbox{if } \theta=1\\
        -\pi_{i}S &\mbox{if } \theta=0
    \end{array}
\right.
$$
Nevertheless, this is from the perspective of a gambler and not the bookie. In order to fully comprehend this scenario, we are obliged to invert this function. The reason for this is that, in reality, we are the bookie and, much like the bookie, we provide our own probability judgements of an event's outcomes $\pi$ - which is referred to as the 'price' above. This, consequently, is sometimes referred as being a bookie with 'nature'. Regardless, below we find the same gain function $g_\theta$ inverted.
$$
g_{\theta} = \left\{
    \begin{array}{ll}
        -(1-\pi_{i})S &\mbox{if } \theta=1\\
          \pi_{i}S &\mbox{if } \theta=0
    \end{array}
\right.
$$
Now that we have inverted our perspective, we can consider a more complex scenario but with less fuddled language. 

We decide to place a series of separate bets on $3$, disjoint, binary events -denoted $\theta_{1}$, $\theta_{2}$, and $\theta_3$ - and a corresponding stake $S_\theta$ for each event, forming a vector $s=(S_{\theta_{1}}, S_{\theta_{2}}, S_{\theta_{3}})$. We can likewise consider the probability of each event $\pi_{\theta_{i}}$, and assign the following probabilities $\pi_{\theta_{1}}$, $\pi_{\theta_{2}}$, and $\pi_{\theta_{3}}$:
```{r}
pi1 <- .2
pi2 <- .8
pi3 <- .1
```
Because the events are mutually exclusive events, we can conveniently construct our probabilities in an $n\times n$ square matrix $R$ to track all possible events and the outcomes therein. Intuitively, this theoretically means that each row, column outcome should meet the condition $\pi_{\theta_{ij}}=\pi_{\theta_{ij}}+\pi_{\theta_{ij}}$. 
```{r}
R <- matrix(
 c(1 - pi1,   - pi2,   - pi3,
     - pi1, 1 - pi2,   - pi3,
     - pi1,   - pi2, 1 - pi3),
            nrow = 3)
R
```
Accordingly, we can then use the vectors of stakes $s$ multiplied with our $n\times n$ matrix $R$ to solve for the vector of gains $g_{\theta}$, such that $g_{\theta}=Rs$.

However, it is now critical to bring to light the following for the reader who has not yet noticed: our set of bets do not conform to Kolmogorov’s probability axiom $P(A\cap, ...,\cap A) = 1$ for a mutually exclusive set of events. As the sum of the bets is
```{r, echo = TRUE}
sum(pi1 + pi2 + pi3)
```
it clearly implies that $\pi_{\theta_{1}}+\pi_{\theta_{2}}+\pi_{\theta_{3}}\neq0$. Consequently, the matrix $R$ is invertible. Technically speaking, and using a more commonly adopted notation, there is a scalar $x$ that gives the desired output $b = Ax$. As a corollary, this also means that the determinant of the matrix $R$ is $|{R}| \neq 0$, as evidence by
```{r, echo = TRUE}
det(R)
```
Thus, we are obliged to prove whether our matrix $R$ does conform to this condition, if $I = R^{-1}R$. Nevertheless, we must first digress by explaining the properties of an identity matrix $I$, for those who are unfamiliar. 

An identity matrix $I$ is where all the diagonals of a matrix are $1$'s, with the rest of the elements only containing $0$'s. It is, broadly speaking, a convenient tool used for matrix manipulation. Below is an example of an $n\times n$ identity matrix $I$:
```{r, echo = FALSE}
I <- diag(c(1, 1, 1))
I
```
It is, therefore, a literal 'do nothing' matrix, as is clearly evidence by solving the following,
```{r, echo = TRUE}
b <- c(1, 2, 3)
I%*%b
```
Evidently, $I\times b$ does not change the initial values of $b$, such that $I\times b = b$. This, then, brings us back to the focal point of this discussion. 

If $R$ is indeed an invertible matrix, we will be able to recreate the same identity matrix $I$ as above, where $I = R^{-1}R$.
```{r, echo = TRUE}
round(inv(R)%*%R, digits = 5)
```

The result of applying $R^{-1}R$ clearly mirrors the identity matrix example given prior. As already alluded to, proving $I = R^{-1}R$ can be thought as a soft proof for whether any combination of linear equations is solvable if, for example, we wanted to bring $R$ to the other side of the equation and solve for $s$. Consequently, if it is proven, this makes it possible to solve for the vector of stakes $s$, by assuming them to be an unknown, and, in turn, by assigning a vector of desired gains to $g_{\theta}$. For instance, if we assign the vector $g_{\theta}$ the following positive, desired gains,
```{r, echo = TRUE}
g_theta <- c(400, 400, 400)
g_theta
```
it is then possible to solve for the linear transformation of $s$, such that we are guaranteed the desired outcomes initially assigned to $g_\theta$ above. First, multiply each side of the equation by $R^{-1}$, so that $R^{-1}Rs=R^{-1}g$, thus $I\times s = s =R^{-1}g$, and, ultimately $s = R^{-1}g$. Accordingly, we can solve for a vector of stakes $s$ that guarantees a net positive gain for each event,
```{r, echo = TRUE}
s_solved <- inv(R)%*%g_theta
s_solved
```
and, in reverse, as is obvious from the above,
```{r, echo = TRUE}
R%*%s_solved
```
we get our desired gains $g_\theta$. 

To conclude, it is clear that, if the probability matrix $R$ meets the invertible condition as has been described here, such that $I = R^{-1}R$, we can solve for an unknown vector of stakes $s$ by equating it with the product of a linear combination of desired gains $g_{\theta}$ and an associated probability matrix $R$. In technical language, this means that when $R$ is an invertible matrix, the solution to $Rs=g_{\theta}$ is simply $s=R^{-1}g$.