---
title: "DT_proof"
author: "Josh Soboil"
date: "13/09/2020"
output: 
 html_document:
  theme: readable
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(matlib)
```

##  An incoherent (Dutch Book) series of bets:
Suppose we are asked to assign the probabilities for three, mutually exclusive events. Let $E_{1}$, $E_{2}$, $E_{3}$ be the disjoint events and $p_{1}$, $p_{2}$, and $p_3$ the probabilities we associate with each event. More simply, let $p_{i}$ simply be the probability of event $E_{i}$ occurring:
```{r, echo = TRUE}
p1 <- .3
p2 <- .1
p3 <- .4
```
If we consider the total winning stakes as unknown, we can construct our probabilities in an $n\times n$ matrix $R$. Note that we construct the matrix in this way to account for a positive gain for the outcome of event $E_{i}$, and a negative gain for the other events of $E_{i}$ if the better placed a stake $s$ on all possible bets.
```{r, echo = TRUE}
R <- matrix(c(1 - p1,   - p2,   - p3,
                - p1, 1 - p2,   - p3,
                - p1,   - p2, 1 - p3), 
            nrow = 3)
R
```
Thus, the total gains $G$ is the difference between the sum of the set of stakes $s$ and the sum of the $n$ stakes $s$ paid to the bookkeeper of the bet, i.e. the sum of negative bets, such that $G_{n} = S_{n} - \sum_{1}^{n}p_{i}S{n}$.

Let us now give some numerical stake $s$ to each possible event, according to what we see fit with respect to the probability of each event,
```{r, echo = TRUE}
s <- c(5, 2, 7)
s
```

Now that we have a set $p_{i}$ of probabilities corresponding to a set $s$ of stakes, we can solve a system of linear equations which will provide a corresponding Gain $G$, the product of $Rs$, such that $G = Rs$.
```{r, echo = TRUE}
G <- solve(a = R, b = s)
G
```
Our net gains $G$ subsequently equal:
```{r, echo = TRUE}
round(sum(G))
```
We do seem to be very lucky. No matter the outcome, nor the size of the bet, we have managed to obtain  a net positive gain $G$ of $70$. It does not matter what we do, we will _always_ have a positive net gain value. How could we always win no matter what? Sadly, it is not because we are clairvoyant.

The result is due to the fact that our initial probability judgements of the $3$ disjoint events were _incoherent_, according to Kolmogorov’s probability axiom $P(A\cap, ...,\cap A) = 1$. Critically, the set of probabilities that we assigned to each event evidently did not sum to a total of $1$:
```{r, echo = TRUE}
sum(p1 + p2 + p3)
```
This, in turn, means that our matrix $R$ of event probabilities is invertible,
```{r}
matlib::inv(R)
```
Simply, an $n\times n$ identity matrix $I$ can be formed from the multiplication of the $n\times n$ matrix $R$ with its inverse $R^{-1}$. An identity matrix is a matrix where all diagonals of a matrix are $1$'s, and the rest of the elements are $0$'s. For a clear example, below is an $n\times n$ identity matrix $I$:
```{r, echo = FALSE}
I <- diag(c(1, 1, 1))
I
```
Consequently, if our set of event probabilities, represented by the matrix $R$, conform to the above specific matrix behaviour, where we can obtain $I = R^{-1}R$, we can confirm $R$ to be invertible by multiplying the original matrix by its inverse as follows:
```{r, echo = TRUE}
I <- round(inv(R)%*%R)
I
```
This behaviour additionally makes the determinant of the matrix $|{R}| > 0$:
```{r, echo = TRUE}
det(R)
```
and, as a consequence, our net gain $\sum_{1}^{n}G_{i}$ is a constantly positive.

Moreover, and critically important, if our event probabilities are incoherent, and thus lead to an invertible matrix as described above, we can solve for an unknown set of stakes $s$ by equating the set to be the product of a desirable set of knowable gains $G_{n}$ with the corresponding event probabilities $R_{i}$. For instance, let's say we want a definite net gain $G_{n}$ for each series of bets $R_{i}$ as follows:
```{r, echo = TRUE}
G <- c(3, 1, 4)
G
```
We can then solve for $s$, through setting $s = R^{-1}G$:
```{r}
s_solved <- solve(a = inv(R), b = G)
s_solved
```

Intuitively speaking, we can interpret this in a geometric sense: the total 'area' of our set of disjoint probabilities fails to induce a single, whole sample space, where $P(A\cap, ...,\cap A) \neq 1$. In order to be coherent, we are therefore obliged to constrain the total sum area of our independent probabilities to be $= 1$, such that $P(A\cap, ...,\cap A) = 1$. We provide a proof for this below.

## A coherent series of bets:
In contrast, here we will assign probability judgements that conform to Kolmogorov's axiom $P(A\cap, ...,\cap A) = 1$. Note the sum of probabilities $p_{i}$ now equal $1$.
```{r A coherent bet, echo = TRUE}
p1 <- .3
p2 <- .3
p3 <- .4

sum(p1 + p2 + p3)
```
Again, we can conveniently put these event probabilities in the form of a matrix $R$ to present all possible joint positive and negative betting outcomes, while representing all disjoint positive event outcomes.
```{r, echo = TRUE}
R <- matrix(c(1 - p1,   - p2,   - p3,
                - p1, 1 - p2,   - p3,
                - p1,   - p2, 1 - p3), 
            nrow = 3)
R
```
However, because we ensured that $P(A\cap, ...,\cap A) = 1$, where all disjoint events sum to a probability of $1$, the matrix $R$ is now not invertible. An identity matrix $I$ cannot, therefore, be formed, where $I = R^{-1}R$, and is evident from the below calculation:
```{r, echo = TRUE}
I <- round(R^-1%*%R)
I
```
Moreover, now that the system is coherent, it will always result in a net gain of $0$ for any arbitrary series of stakes $s$:
```{r, echo = TRUE}
s <- sample(x = 1:10000, size = 3, replace = TRUE)
s
```
Accordingly, the gains for each bet equal:
```{r, echo = TRUE}
G <- R%*%s
G
```
Decisively, and despite the relatively high stakes on each bet compared to the previous ones, notice the net overall gains maintain to be $0$:
```{r, echo = TRUE}
round(sum(G))
```